---
title: "Homework 6"
author: "Kallan Roan"
date: "2025-11-23"
output: github_document
---

Load libraries

```{r}
library(tidyverse)
library(p8105.datasets)
library(modelr)
```

Set seed

```{r}
set.seed(1)
```


Preset settings

```{r}
knitr::opts_chunk$set(
  fig.width = 9,
  fig.asp = .95,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Import data and clean data to include criteria indicated below:
Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv") |>
  janitor::clean_names() |> 
  mutate(
    city_state = str_c(city, ", ", state),
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age)
  ) |> 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
    ) |> 
  select(resolved, victim_age, victim_race, victim_sex, city_state)

```

Logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors for Baltimore, MD.

```{r}
balt_fit_logistic = 
  homicide_df |> 
  filter(city_state == "Baltimore, MD") |> 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = _, family = binomial())

balt_fit_logistic |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    OR = exp(estimate),
    conf_lower = exp(conf.low),
    conf_upper = exp(conf.high)
  ) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, conf_lower, conf_upper) |> 
  knitr::kable(digits = 3)
```

Logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors for all cities.

Create function to run logistic regression

```{r}
homicide_glm = function(df) {
  
  glm(resolved ~ victim_age + victim_sex + victim_race, data = df, family = binomial())

}
```

Run logistic regression for all cities

```{r}
fit_logistic_all = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    fits = map(data, homicide_glm),
    results = map(fits, \(df) broom::tidy(df, conf.int = TRUE))
  ) |> 
  select(city_state, results) |> 
  unnest(results) |> 
  mutate(
    OR = exp(estimate),
    conf_lower = exp(conf.low),
    conf_upper = exp(conf.high)
  ) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, conf_lower, conf_upper)

fit_logistic_all |> 
  knitr::kable(digits = 3)

```

Create plot that shows OR's and Ci's for each city

```{r}
fit_logistic_all |> 
  mutate(
    city_state = fct_reorder(city_state, OR)
  ) |> 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Odds Ratios of Homicides comparing Males vs Females",
    x = "Location",
    y = "OR"
  ) 
```

The majority of cities (41 of 47 total cities) have odds ratios that are less than 1, meaning that for most cities, the odds of homicides are greater among females than males. For some cities, the odds ratios are significant, since the confidence intervals do not include the null value of 1. On the other hand, only 6 cities have odds ratios greater than 1, meaning that the odds of homicides is greater among males vs females. However, the odds are not significant, since the confidence intervals for all of those odds ratios include the null value of 1. Therefore, most U.S cities follow the trend that females have a higher odds of being vicims of homicide compared to males. 

Albuquerque, NM has the highest, but insignificant odds ratio of 1.767 with the widest 95% confidence interval of (0.825, 3.762). On the other hand, New York, NY has the lowest homicide odds ratio comparing males to females of 0.262 (95% CI: 0.133, 0.485). Chicago, IL has the narrowest 95% confidence interval of (0.336, 0.501) and OR = 0.410

## Problem 2

Load weather dataset

```{r}
data("weather_df")
```

5000 Bootstrap samples. Produce tibbles from tidy and glance.

```{r}
weather_bootstrap_results = 
  weather_df |> 
  bootstrap(n = 5000) |> 
  mutate(
    df = map(strap, as_tibble),
    fits = map(df, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results_1 = map(fits, broom::tidy),
    results_2 = map(fits, broom::glance)
  ) |> 
  select(.id, results_1, results_2)
```

Extract r.squared and ratio

```{r}
weather_bootstrap_results_cleaned = 
  weather_bootstrap_results |> 
  unnest(results_2) |> 
  select(.id, results_1, r.squared) |> 
  unnest(results_1) |> 
  select(.id, term, estimate, r.squared) |> 
  filter(term != "(Intercept)") |> 
  pivot_wider(
    names_from = term, 
    values_from = estimate
  ) |> 
  mutate(
    ratio = tmin / prcp
  ) |> 
  select(.id, r.squared, ratio)
```

Plot of distribution of ratio

```{r}
weather_bootstrap_results_cleaned |> 
  ggplot(aes(x = ratio)) +
  geom_density()
```

The distribution of the ratio of b1 and b2 is centered at around -180, with a range between under -400 to around -100. In addition, the distribution also has a heavy tail extending to the lower values. 

Plot of distribution of r.squared

```{r}
weather_bootstrap_results_cleaned |> 
  ggplot(aes(x = r.squared)) +
  geom_density()
```

The distribution of r.squared is centered at around 0.940, with a range between 0.92 to 0.95. This means that a higher percentage of variation in tmax is explained by variation in tmin and prcp in the model, demonstrating that the model with tmin and prcp as predictors and tmax as the outcome is a good model. In addition, the distribution has a heavy tail extending to the lower values. 


95% confidence intervals for r.squared and ratio

```{r}
weather_bootstrap_results_cleaned |> 
  pivot_longer(
    cols = r.squared:ratio,
    names_to = "term",
    values_to = "estimate"
  ) |> 
  group_by(term) |> 
  summarize(
    ci_lower = quantile(estimate, 0.025),
    ci_upper = quantile(estimate, 0.975)
  ) |> 
  knitr::kable()
```

## Problem 3

Import and clean birthweight.csv

```{r}
birthweight_df =
  read_csv("data/birthweight.csv") |>
  janitor::clean_names() |> 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) |> 
  drop_na()
```

Proposed regression model

My proposed regression model is entirely theory-based. I believe relying on hypothesis will allow for the most clinically accurate prediction of birthweight. I aim to include variables that are related to the baby and the mother, since traits about the baby and the mother can influence birthweight. My chosen model will include main effects of `bhead`, `blength`, `gaweeks`, `smoken`, and `mrace` as predictors of bwt. 

A longer `bhead` and `blength`, representing a baby's head circumference and length at birth, is associated with higher birthweight. Additionally, `gaweeks`, representing gestational age in weeks, can influence birthweight, since premature babies are known to have lower birthweight. `smoken`, representing average number of cigarettes smoked per day during pregnancy, can influence birthweight, since smoking can lead to lower birthweight. Finally, a mother's race (`mrace`) is an important predictor to include in the regression model, since the certain race groups are known to have higher prevalence of low birthweight babies. Therefore, my chosen final regression model will include variables related to both the infant and the mother to produce the best prediction of birthweight. 

Plot of Predicted vs Fitted

```{r}
model_1 = lm(bwt ~  bhead + blength + gaweeks + smoken + mrace, data = birthweight_df)

birthweight_df |> 
  add_predictions(model_1) |> 
  add_residuals(model_1) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "red") +
  labs(
    title = "Fitted vs Residuals",
    x = "Fitted Values",
    y = "Residuals"
  ) 
```

Comparison with other models

```{r}
cv_df =
  crossv_mc(birthweight_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
  
cv_df = 
  cv_df |> 
  mutate(
    model_1 = map(train, \(df) lm(bwt ~ bhead + blength + gaweeks + smoken + mrace, data = df)),
    model_2 = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model_3 = map(train, \(df) lm(bwt ~ bhead + blength + babysex + (bhead * blength * babysex), data = df))
  ) |> 
  mutate(
    rmse_model1 = map2_dbl(model_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model2 = map2_dbl(model_2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model3 = map2_dbl(model_3, test, \(mod, df) rmse(model = mod, data = df))
  )
```

Compare RMSE between models

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```

Based on the comparison between the three models, the 2nd model with length and gestational age as main effects has the has the highest average RMSE values compared to the 1st model and 3rd model, demonstrating that the 2nd model is not the most optimal model to predict infant birthweight. The 1st model using head circumference, length, gestational weeks, smoking, and mother's race as predictors and the third model using head circumference, length, sex, and all interactions as predictors are comparable in predicting birthweight. 
