---
title: "Homework 6"
author: "Kallan Roan"
date: "2025-11-23"
output: github_document
---

Load libraries

```{r}
library(tidyverse)
library(p8105.datasets)
set.seed(1)
```

Preset settings

```{r}
knitr::opts_chunk$set(
  fig.width = 9,
  fig.asp = .95,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Import data and clean data to include criteria indicated below:
Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv") |>
  janitor::clean_names() |> 
  mutate(
    city_state = str_c(city, ", ", state),
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age)
  ) |> 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
    ) |> 
  select(resolved, victim_age, victim_race, victim_sex, city_state)

```

Logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors for Baltimore, MD.

```{r}
balt_fit_logistic = 
  homicide_df |> 
  filter(city_state == "Baltimore, MD") |> 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = _, family = binomial())

balt_fit_logistic |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(
    OR = exp(estimate),
    conf_lower = exp(conf.low),
    conf_upper = exp(conf.high)
  ) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, conf_lower, conf_upper) |> 
  knitr::kable(digits = 3)
```

Logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors for all cities.

Create function to run logistic regression

```{r}
homicide_glm = function(df) {
  
  glm(resolved ~ victim_age + victim_sex + victim_race, data = df, family = binomial())

}
```

Run logistic regression for all cities

```{r}
fit_logistic_all = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    fits = map(data, homicide_glm),
    results = map(fits, \(df) broom::tidy(df, conf.int = TRUE))
  ) |> 
  select(city_state, results) |> 
  unnest(results) |> 
  mutate(
    OR = exp(estimate),
    conf_lower = exp(conf.low),
    conf_upper = exp(conf.high)
  ) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, conf_lower, conf_upper)

fit_logistic_all |> 
  knitr::kable(digits = 3)

```

Create plot that shows OR's and Ci's for each city

```{r}
fit_logistic_all |> 
  mutate(
    city_state = fct_reorder(city_state, OR)
  ) |> 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Odds Ratios of Homicides comparing Males vs Females",
    x = "Location",
    y = "OR"
  ) 
```

COMMENT HERE??????


